Evolving Dynamics: The U.S. Labor Force from 1972 to 2022 and the Enduring Challenges for Women in the Workplace

### 2024-03-15 paper/paper.qmd
rendered pdf
updated results and cleaning based on rohan comments

```{r}
#| echo: false
#| tbl-cap: Women in the workforce survey data

library(kableExtra)

@@

library(ggplot2)

data <- read.csv("../data/analysis_data/full_time_data.csv")

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```

@@

```{r}
#| label: fig-part_time
#| fig-cap: Part Time Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

library(ggplot2)

data <- read.csv("../data/analysis_data/part_time_data.csv")

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```

@fig-part_time highlights the number of men and women engaged in part-time work from 1972 to 2022, showcasing evolving trends in part-time employment, with a noticeable gender dimension. In 1972, part-time work was more common among women (73) than men (41), a trend that persisted and widened over the decades. The gap between male and female part-time workers expanded, with women consistently outnumbering men in part-time positions. Notably, the data illustrates a significant jump in both men and women working part-time in 1994, suggesting a shift in employment patterns or economic conditions that increased part-time work availability or necessity. By 2022, the number of women working part-time reached 219, compared to 126 men, reflecting ongoing gender disparities in part-time employment.

@@

library(ggplot2)

data <- read.csv("../data/analysis_data/unemployed_data.csv")

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```
@fig-unemployed captures the number of unemployed men and women from 1972 to 2022, providing insightful trends into the dynamics of unemployment across genders over five decades. In 1972, the dataset begins with 42 unemployed men compared to 8 unemployed women, indicating a higher male unemployment in absolute numbers. Over the years, both genders experienced fluctuations in unemployment, with notable increases during economic downturns. A significant spike in unemployment for both genders is observed in 2021, with 181 men and 150 women reported as unemployed, possibly reflecting the economic impacts of global events or shifts in the labor market. While the gap between unemployed men and women has varied, the year 2021 shows a notably smaller difference, suggesting a convergence in unemployment rates between genders.

## Biases against hiring women

@@

library(ggplot2)
library(readxl)
library(tidyr)

# Read the data from the Excel file
data <- read_excel("../data/analysis_data/pref_hiring.xlsx")

# Reshape the data to long format for easier plotting with ggplot2
data_long <- pivot_longer(data, cols = -Year, names_to = "Category", values_to = "Count")

@@

@@

library(ggplot2)
library(readxl)
library(tidyr)

@@

# Reshape the data to long format for line graphs and convert 'count' to numeric
fefam_data_long <- fefam_data |>
  pivot_longer(cols = c(strongly_agree, agree, disagree, strongly_disagree),
               names_to = "Response to: better for man to work, woman tend home",
               values_to = "count") |>
  mutate(count = as.numeric(count))

# fefam line graph
ggplot(fefam_data_long, aes(x = year, y = count, group = `Response to: better for man to work, woman tend home`, color = `Response to: better for man to work, woman tend home`)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1200)) +
  labs(title = "Responses to: better for man to work, woman tend home",
       x = "Year",
       y = "Count") +
  theme_minimal()
```

>>>>

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: Women in the workforce survey data
library(tidyverse)
library(kableExtra)

@@

library(tidyverse)

ft_data <- read_csv("../data/analysis_data/full_time_data.csv")

ggplot(ft_data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```

@@

```{r}
#| label: fig-part-time
#| fig-cap: Part Time Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

library(tidyverse)

pt_data <- read_csv("../data/analysis_data/part_time_data.csv")

ggplot(pt_data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```

@fig-part-time highlights the number of men and women engaged in part-time work from 1972 to 2022, showcasing evolving trends in part-time employment, with a noticeable gender dimension. In 1972, part-time work was more common among women (73) than men (41), a trend that persisted and widened over the decades. The gap between male and female part-time workers expanded, with women consistently outnumbering men in part-time positions. Notably, the data illustrates a significant jump in both men and women working part-time in 1994, suggesting a shift in employment patterns or economic conditions that increased part-time work availability or necessity. By 2022, the number of women working part-time reached 219, compared to 126 men, reflecting ongoing gender disparities in part-time employment.

@@

library(tidyverse)

u_data <- read_csv("../data/analysis_data/unemployed_data.csv")

ggplot(u_data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```

@fig-unemployed captures the number of unemployed men and women from 1972 to 2022, providing insightful trends into the dynamics of unemployment across genders over five decades. In 1972, the dataset begins with 42 unemployed men compared to 8 unemployed women, indicating a higher male unemployment in absolute numbers. Over the years, both genders experienced fluctuations in unemployment, with notable increases during economic downturns. A significant spike in unemployment for both genders is observed in 2021, with 181 men and 150 women reported as unemployed, possibly reflecting the economic impacts of global events or shifts in the labor market. While the gap between unemployed men and women has varied, the year 2021 shows a notably smaller difference, suggesting a convergence in unemployment rates between genders.

@@

library(tidyverse)
library(ggplot2)
library(readxl)
library(tidyr)

# Read the data from the Excel file
h_data <- read_excel("../data/analysis_data/pref_hiring.xlsx")

# Reshape the data to long format for easier plotting with ggplot2
data_long <- pivot_longer(h_data, cols = -Year, names_to = "Category", values_to = "Count")

@@

## Biases against hiring women
@fig-prefhiring captures public opinion on hiring preferences across four categories: "Strongly favor," "Not strongly favor," "Not strongly oppose," and "Strongly oppose" from 1996 to 2006. The data reveals a fluctuating yet insightful perspective on preferences towards hiring practices over the years. In 1996, 139 individuals strongly favored certain hiring preferences, while a significant 453 strongly opposed them, indicating a polarized view on the matter. Over the following years, the number of people strongly favoring hiring preferences sees variation, peaking in 2006 with 204 individuals. Conversely, the "Strongly oppose" category also witnesses fluctuations, decreasing to 320 in 2000, then varying in subsequent years, and peaking again at 372 in 2006.

@@

library(tidyverse)
library(ggplot2)
library(readxl)
library(tidyr)

@@

# Reshape the data to long format for line graphs and convert 'count' to numeric
fefam_data_long <- fefam_data |>
  pivot_longer(cols = c(strongly_agree, agree, disagree, strongly_disagree),
               names_to = "Responses",
               values_to = "count") |>
  mutate(count = as.numeric(count))

# fefam line graph
ggplot(fefam_data_long, aes(x = year, y = count, group = `Responses`, color = `Responses`)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1200)) +
  labs(title = "Responses to: better for man to work, woman tend home",
       x = "Year",
       y = "Count") +
  theme_minimal()
```
